---
title: "Database Code Along Session"
format: html
editor: visual
---

## Load Packages
```{r}

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, RPostgres, DBI, dbplyr)

```

## Connect to `ribbitr` Database - `usethis::edit_r_environ()`
```{r}

ribbitr_connection <- dbConnect(drv = dbDriver("Postgres"),
                                dbname = Sys.getenv("aws_dbname"),
                                host = Sys.getenv("aws_host"),
                                port = Sys.getenv("aws_port"),
                                user = Sys.getenv("aws_user"),
                                password = Sys.getenv("aws_password"))


```

## Set `search_path` for `survey_data` Schema
```{r}

search_path <- "set search_path = 'survey_data';"

dbExecute(ribbitr_connection, search_path)

```

## Sample Query with `DBI` Package

### `location` Table
```{r}

sql <- "SELECT * FROM location;"

location <- dbGetQuery(ribbitr_connection, sql)


```

### Join `location` and `region` Tables Together
```{r}

sql_loc_reg <- "SELECT l.location, r.region
                FROM location l
                JOIN region r on l.location_id = r.location_id;"

loc_reg <- dbGetQuery(ribbitr_connection, sql_loc_reg)

```

### Join to `capture` Level Data - Join tables first then select columns
```{r}

# use `*` to show difference
cap_sql <- "SELECT l.location, r.region, s.site, v.date, v.survey_time, s2.start_time, s2.end_time, c.species_capture
            FROM location l
            JOIN region r on l.location_id = r.location_id
            JOIN site s on r.region_id = s.region_id
            JOIN visit v on s.site_id = v.site_id
            JOIN survey s2 on v.visit_id = s2.visit_id
            JOIN capture c on s2.survey_id = c.survey_id;"

clean_data <- dbGetQuery(ribbitr_connection, cap_sql)

```

### Join to `capture` Level Data - Add `WHERE` Clause
```{r}

# use `*` to show difference
penn_sql <- "SELECT l.location, r.region, s.site, v.date, v.survey_time, s2.start_time, s2.end_time, c.species_capture
             FROM location l
             JOIN region r on l.location_id = r.location_id
             JOIN site s on r.region_id = s.region_id
             JOIN visit v on s.site_id = v.site_id
             JOIN survey s2 on v.visit_id = s2.visit_id
             JOIN capture c on s2.survey_id = c.survey_id
             WHERE r.region = 'pennsylvania' AND v.survey_time = 'day';"

penn_data <- dbGetQuery(ribbitr_connection, penn_sql)
```

## Sample Query with `dbplyr` Package

### `location` Table
```{r}
# Storing a database version table in memory of the `location` table
location_table <- tbl(ribbitr_connection, "location")

# retrieve data
location_table <- tbl(ribbitr_connection, "location") %>% 
  collect()

# Display SQL query
tbl(ribbitr_connection, "location") %>%
  show_query()

```

### Join `location` and `region` Table
```{r}
loc_reg <- tbl(ribbitr_connection, "location") %>%
  left_join(tbl(ribbitr_connection, "region"), by = c("location_id")) %>%
  select(c(location, region)) %>%
  collect()

```

### Join to `capture` Level Data
```{r}

db_data <- tbl(ribbitr_connection, "location") %>%
  left_join(tbl(ribbitr_connection, "region"), by = c("location_id")) %>%
  left_join(tbl(ribbitr_connection, "site"), by = c("region_id")) %>%
  left_join(tbl(ribbitr_connection, "visit"), by = c("site_id")) %>%
  left_join(tbl(ribbitr_connection, "survey"), by = c("visit_id")) %>%
  left_join(tbl(ribbitr_connection, "capture"), by = c("survey_id")) %>%
  select(c(location, region, site, date, start_time, end_time, survey_time, duration_minutes,
           species_capture, body_mass_g, svl_mm, life_stage, sex)) %>% 
  filter(region == "pennsylvania", survey_time == "day") %>% 
  collect()


```


